### 2016-09-28 rebuttal.tex

\begin{document}

\includepdf{coverletter.pdf}

\section{Comments from Editor}

\begin{questions}
\question I found the \url{http://live.sympy.org/} site very convenient for trying things
out in SymPy, but it is only mentioned in the supplement, which might be
overlooked by readers. I suggest moving that section into the main paper, as
it provides a great way to play along with the examples while reading the
paper.
\begin{solution}
\label{editorpoint1}
Section (previously) 9 from the supplement was edited and moved to the introduction:
\begin{quote}
All examples could be tested on the \href{http://live.sympy.org}{SymPy
Live} instance, that is an online Python shell, which uses the Google
App Engine to execute SymPy code.
\end{quote}
\end{solution}
\question The Basic Usage section omits what I think is an important point:
how does one distinguish between evaluating exp(1) in Python and exp(1) in
SymPy? In other words, how are symbolic constants specified?
\begin{solution}
  Both Python standard library and SymPy have exp callable, so it depends
  which symbols are imported in your global namespace: the \texttt{exp}
  function from the \texttt{math} module or the \texttt{exp} class from SymPy.
  In the later case a special symbolic constant will be produced.

@@

We have added a sentence to the section~\ref{P-sec:core}, that shows symbolic
constants as an example of leaf nodes:
\begin{quote}
Symbols or symbolic constants, like $e$ or $\pi$, are other examples of
leaf nodes.
\begin{verbatim}
>>> exp(1)
E
>>> exp(1).args
()
>>> x.args
()
\end{verbatim}
\end{quote}
\end{solution}
\question The first
thing I tried after seeing Table 2 (simplication functions) did not work:

\begin{verbatim}
>>> trigsimp (exp( Matrix(2, 2, [0, -y, y, 0])) )
\end{verbatim}
fails to recognize cos and
sin. Am I expecting too much?
\begin{solution}
  This indeed does not work. The implementation in \verb|trigsimp| primarily
  thus far has focused on transforming trigonometric functions into other
  trigonometric functions. The ability to simplify complex exponentials is
  something that we would like to work.
  \href{https://github.com/sympy/sympy/issues/11459}{Issue 11459} in our
  public issue tracker tracks this feature.
\end{solution}
\question Section 3.6: are eigenvalues and singular
values included? If not, why not?
\begin{solution}
Yes, they are included.  The second example in that section addresses the computation of
eigenvalues.  Singular values can be computed with
\texttt{A.singular\_values()}. We have not included this example, as the
singular value expressions for our example matrix \texttt{A} are quite large,
but we have added ``singular values'' to the list of features in the second
paragraph of the section.
\end{solution}
\question Like one of the referees, I expected to
see MPFR mentioned in Section 4.1, at least to mention the pros and cons and
say why it isn't used.
\begin{solution}

  The major advantage is that mpmath is a pure Python library. It's slower
  than MPFR, but not too much to evaluate any reasonably complicated single
  expression at interactive speed, which covers most use cases in SymPy. We
  have added a sentence about this to the numerics section.

We mention this at the beginning of the section. Another advantage is that
mpmath is BSD licensed, like SymPy, which is important for the scientific
Python ecosystem (MPFR is licensed under the LGPL).
\end{solution}
\question Section 4.1: please state whether the syntax for
functions in mpmath is identical, or not, to that in Sympy.
\begin{solution}
The mpmath library is not a SymPy submodule, but a separate library,
which we have indicated in the text. Most function names in mpmath are the same as SymPy, but it is not true in every case. We have
added an example to the text showing an instance where they diverge
(summations).
\end{solution}
\question Unless I have
missed something, a weakness of the paper is that speed is mentioned only
twice, on lines 496 and 666. My assumption is that SymPy is slow compared with
its commercial competitors. Please comment on the speed issue, and not just in
the conclusions.
\begin{solution}
A new paragraph titled ``Speed'' was added to the manuscript, which discusses the
speed issue and lays out a path forward. The conclusion was simplified.
\end{solution}
\question Editors are needed for [20].
\begin{solution}
We have done this.
\end{solution}
\end{questions}

### 2016-09-28 projects_that_depend_on_sympy.tex
Fix the width of the projects that depend on SymPy table
Add citation for SymPy Gamma
Move citations to the text in the projects table

\begin{longtable}[htbc]{>{\raggedright}p{0.2\linewidth}p{0.8\linewidth}}
\caption{Selected projects that depend on SymPy.\label{projects-table}}\\
\toprule
\textbf{Project name} & \textbf{Description} \\
\midrule

\href{http://sympygamma.com/}{\textbf{SymPy Gamma}} & An open source
  analog of the Wolfram|Alpha.  See more in the supplementary material. \\

\href{http://cadabra.science/index.html}{\textbf{Cadabra}}~\cite{Peeters2007cadabra} &
  CAS designed specifically for the resolution of problems
  encountered in field theory. \\

\href{https://github.com/cbm755/octsympy}{\textbf{GNU Octave Symbolic Package}}~\cite{OctSymPy} &
  An implementation of a symbolic toolbox for Octave using SymPy. \\

\href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}~\cite{SymPy.jl} &
  Provides a Julia interface to SymPy using PyCall. \\

\href{https://mathics.github.io/}{\textbf{Mathics}}~\cite{Mathics} &
  A free, online CAS featuring Mathematica compatible
  syntax and functions. \\

\href{http://mathpix.com/}{\textbf{Mathpix}}~\cite{Mathpix} & An iOS App, that detects handwritten math as input, and uses
  SymPy Gamma to evaluate the math input and generate the relevant
  steps to solve the problem. \\

\href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}}~\cite{diankov2010ikfast} &
  IKFast is a robot kinematics compiler provided by
  \href{http://openrave.org/}{OpenRAVE}. \\

\href{http://www.sagemath.org/}{\textbf{SageMath}}~\cite{sagemath} &
  A free open-source mathematics software system, builds on top of many
  existing open-source packages, including SymPy. \\

\href{http://www.pydy.org/}{\textbf{PyDy}}~\cite{gede2013constrained} & Multibody Dynamics with
  Python. \\

\href{https://github.com/brombo/galgebra}{\textbf{galgebra}}~\cite{galgebra} &
  A module for geometric algebra (previously \texttt{sympy.galgebra}). \\

\href{http://yt-project.org/}{\textbf{yt}}~\cite{2011ApJS..192....9T} & Python package for
  analyzing and visualizing volumetric data. \\

\href{http://sfepy.org/}{\textbf{SfePy}}~\cite{cimrman2014sfepy} &
  A Python package for solving partial
  differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
  method~\cite{Zienkiewicz2013FEM}. \\

\href{http://quameon.sourceforge.net/}{\textbf{Quameon}}~\cite{quameon} & Quantum
  Monte Carlo in Python. \\

\href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}~\cite{lcapy} &
  Experimental Python package for teaching linear circuit analysis. \\
\bottomrule
\end{longtable}

>>>>

\begin{longtable}[htbc]{>{\raggedright}p{0.2\linewidth}p{0.74\linewidth}}
\caption{Selected projects that depend on SymPy.\label{projects-table}}\\
\toprule
\textbf{Project name} & \textbf{Description} \\
\midrule

\href{http://sympygamma.com/}{\textbf{SymPy Gamma}} & An open source
  analog of Wolfram|Alpha that uses SymPy~\cite{SymPyGamma}.  See more in the supplementary material. \\

\href{http://cadabra.science/index.html}{\textbf{Cadabra}} &
  CAS designed specifically for the resolution of problems
  encountered in field theory~\cite{Peeters2007cadabra}. \\

\href{https://github.com/cbm755/octsympy}{\textbf{GNU Octave Symbolic Package}} &
  An implementation of a symbolic toolbox for Octave using SymPy~\cite{OctSymPy}. \\

\href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}} &
  Provides a Julia interface to SymPy using PyCall~\cite{SymPy.jl}. \\

\href{https://mathics.github.io/}{\textbf{Mathics}} &
  A free, online CAS featuring Mathematica compatible
  syntax and functions~\cite{Mathics}. \\

\href{http://mathpix.com/}{\textbf{Mathpix}} & An iOS App, that detects handwritten math as input, and uses
  SymPy Gamma to evaluate the math input and generate the relevant
  steps to solve the problem~\cite{Mathpix}. \\

\href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}} &
  IKFast is a robot kinematics compiler provided by
  \href{http://openrave.org/}{OpenRAVE}~\cite{diankov2010ikfast}. \\

\href{http://www.sagemath.org/}{\textbf{SageMath}} &
  A free open-source mathematics software system, builds on top of many
  existing open-source packages, including SymPy~\cite{sagemath}. \\

\href{http://www.pydy.org/}{\textbf{PyDy}} & Multibody Dynamics with
  Python~\cite{gede2013constrained}. \\

\href{https://github.com/brombo/galgebra}{\textbf{galgebra}} &
  A module for geometric algebra (previously \texttt{sympy.galgebra})~\cite{galgebra}. \\

\href{http://yt-project.org/}{\textbf{yt}} & Python package for
  analyzing and visualizing volumetric data~\cite{2011ApJS..192....9T}. \\

\href{http://sfepy.org/}{\textbf{SfePy}} &
  A Python package for solving partial
  differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
  method~\cite{Zienkiewicz2013FEM,cimrman2014sfepy}. \\

\href{http://quameon.sourceforge.net/}{\textbf{Quameon}} & Quantum
  Monte Carlo in Python~\cite{quameon}. \\

\href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}} &
  Experimental Python package for teaching linear circuit analysis~\cite{lcapy}. \\
\bottomrule
\end{longtable}
